---
title: "RBarad_HW2"
output: html_document
date: "2023-09-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.ritna = 3)
```

## Import Libraries

```{r import_libraries}
library(tidycensus)
library(tidyverse)
library(kableExtra)
library(sf)
```

# Set options and create color palette

```{r set_options, echo=FALSE}

options(scipen=999) #
options(tigris_class = "sf")

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#ffffd4","#fed98e","#fe9929","#d95f0e","#993404")

```

# Get Census Data

## Set ACS variables

```{r set_variables}

acs_variables2017 <- c(
                   'B25119_001', #Median Household Income
                   'B25058_001', #Rent
                   'B08130_001', #Total Working Population
                   'B08130_016', #Total Population Commuting on Public Transit
                   'B08126_014', #Public Administration Workforce
                   'B17009_001') #Total Population

acs_variables2009 <- c(
  'B25119_001', #Median Household Income
  'B25058_001', #Rent
  'B08130_001', #Total Working Population
  'B08130_016', #Total Population Commuting on Public Transit
  'C24030_028', #Public Administration Workforce Male
  'C24030_055', #Public Administration Workforce Female
  'B17009_001') #Total Population
```


## Import 2009 Census Data

```{r import_2009_data}
dc2009tracts <- get_acs('tract',
                        variables = acs_variables2009,
                        year = 2009,
                        state = 'DC',
                        geometry = TRUE,
                        output = 'wide') %>%
  rename('median_income'='B25119_001E',
         'median_rent'='B25058_001E',
         'working_pop' = 'B08130_001E',
         'transit_pop' = 'B08130_016E',
         'male_public_pop' = 'C24030_028E',
         'female_public_pop' = 'C24030_055E',
         'total_pop' = 'B17009_001E') %>%
  st_transform('EPSG:2248') %>%
  mutate(pct_transit = (transit_pop / working_pop) * 100,
         public_pop = male_public_pop + female_public_pop,
         pct_public = (public_pop / working_pop) * 100,
         year = '2009') %>%
  dplyr::select(-NAME, -starts_with("B0"), -starts_with("B1"), -starts_with("B2"),-starts_with("C2"), -male_public_pop, -female_public_pop)
```

## Import 2017 Census Data

```{r import_2017_data}
dc2017tracts <- get_acs('tract',
                        variables = acs_variables2017,
                        year = 2017,
                        state = 'DC',
                        geometry = TRUE,
                        output = 'wide') %>%
  rename('median_income'='B25119_001E',
         'median_rent'='B25058_001E',
         'working_pop' = 'B08130_001E',
         'transit_pop' = 'B08130_016E',
         'public_pop' = 'B08126_014E',
         'total_pop' = 'B17009_001E',) %>%
  st_transform('EPSG:2248') %>%
  mutate(pct_transit = (transit_pop / working_pop) * 100,
         pct_public = (public_pop / working_pop) * 100,
         year = '2017') %>%
  dplyr::select(-NAME, -starts_with("B0"), -starts_with("B1"), -starts_with("B2"))
```

# Procees DC Metro Data

## Get DC Metro Data
```{r get_metro_data}
dc_boundary <- st_union(dc2017tracts) #Create union of all census tracts in DC

dcmetro <- st_read('https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/51/query?outFields=*&where=1%3D1&f=geojson') %>%
  st_transform('EPSG:2248') #Download DC Metro points and set projection
dcmetro <- dcmetro[st_intersects(dcmetro, dc_boundary) %>% lengths > 0, ] #Intersect DC metro points with DC boundary
```

## Buffer Metro Points

```{r buffer_metro}
metro_buffer <- st_union(st_buffer(dcmetro,1320))
```

## Identify TOD and Non-TOD
```{r identify_TOD}
tod2017 <- dc2017tracts[st_intersects(dc2017tracts, metro_buffer) %>% lengths > 0, ] %>%
  mutate(type='TOD')

non_tod2017 <- dc2017tracts[st_disjoint(dc2017tracts, metro_buffer) %>% lengths > 0, ] %>%
  mutate(type='Non-TOD')

tod2009 <- dc2009tracts[st_intersects(dc2009tracts, metro_buffer) %>% lengths > 0, ] %>%
  mutate(type='TOD')

non_tod2009 <- dc2009tracts[st_disjoint(dc2009tracts, metro_buffer) %>% lengths > 0, ] %>%
  mutate(type='Non-TOD')

tod_dissolve <- st_union(tod2009)

all_census_data <- rbind(non_tod2017,tod2017,tod2009,non_tod2009)
```

# Get Other Data for Map

```{r get_other_data}
dcmetrolines <- st_read('https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/58/query?outFields=*&where=1%3D1&f=geojson') %>%
  st_transform('EPSG:2248') #Download DC Metro points and set projection
dcmetrolines <- st_intersection(dcmetrolines,dc_boundary)

dcwater <- st_read('https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Planimetrics_2015/MapServer/14/query?outFields=*&where=1%3D1&f=geojson') %>%
  filter(DESCRIPTION != 'Pond' & DESCRIPTION != 'Pool')
```

# Make Maps

```{r make_tod_map}
ggplot()+
  geom_sf(data = all_census_data, aes(fill = type))+
  geom_sf(data = metro_buffer,fill='transparent',color='red')+
  facet_wrap(~year)+
  mapTheme()
```

## Map1

```{r make_income_map}

all_census_data <- all_census_data %>% 
  mutate(income_class = cut(median_income, breaks = c(0, 25000, 50000, 75000, 100000, max(all_census_data$median_income, na.rm=TRUE))))

ggplot()+
  geom_sf(data = all_census_data, aes(fill = income_class))+
  geom_sf(data = dcwater,fill='#5ee4ff',linewidth=0)+
  #geom_sf(data = dcmetrolines,linewidth=0.5)+
  geom_sf(data = tod_dissolve,fill='transparent',color='#2dc11e',linewidth=1)+
  geom_sf(data = dcmetro)+
  facet_wrap(~year)+
  scale_fill_manual(values = palette5,
                    name = "Median Income")+
  mapTheme()
```

## Map2

```{r make_rent_map}
ggplot()+
  geom_sf(data = all_census_data, aes(fill = q5(median_rent)))+
  geom_sf(data = tod_dissolve,fill='transparent',color='red',linewidth=1)+
  geom_sf(data = dcmetro)+
  facet_wrap(~year)+
  scale_fill_manual(values = palette5,
                    name = "Median Rent\n(Quintile Breaks)")+
  mapTheme()
```

## Map3

```{r make_transit_usage_map}

ggplot()+
  geom_sf(data = all_census_data, aes(fill = q5(pct_transit)))+
  geom_sf(data = tod_dissolve,fill='transparent',color='red',linewidth=1)+
  geom_sf(data = dcmetro)+
  facet_wrap(~year)+
  scale_fill_manual(values = palette5,
                    name = "Percent Transit\n(Quintile Breaks)")+
  mapTheme()
```

## Map4

```{r make_publicadmin_map}
ggplot()+
  geom_sf(data = all_census_data, aes(fill = q5(pct_public)))+
  geom_sf(data = tod_dissolve,fill='transparent',color='red',linewidth=1)+
  geom_sf(data = dcmetro)+
  facet_wrap(~year)+
  scale_fill_manual(values = palette5,
                    name = "Percent Public\n(Quintile Breaks)")+
  mapTheme()
```


## Lapply Maps - this the map code from Nohman
```{r make_lapply_maps}
maps <- lapply(c("median_income", "median_rent", "pct_transit", "pct_public"), function(i) {  
  ggplot()+  
    geom_sf(data = all_census_data, aes(fill = q5(all_census_data[[i]])))+  
    geom_sf(data = metro_buffer,fill='transparent',color='red')+  
    facet_wrap(~year)+  
    scale_fill_manual(values = palette5,                    
                   name = "Median Income\n(Quintile Breaks)")+  
  mapTheme()})

print(maps)
```

# Groupped Bar Plots

```{r groupped_bar_plot}
census_data_summary <- 
  st_drop_geometry(all_census_data) %>%
  group_by(year, type) %>%
  summarize(median_income = mean(median_income, na.rm = T),
            median_rent = mean(median_rent, na.rm = T),
            percent_transit = mean(pct_transit, na.rm = T),
            percent_public = mean(pct_public, na.rm = T))

census_data_summary %>%
  gather(Variable, Value, -year, -type) %>%
  ggplot(aes(year, Value, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, scales = "free", ncol=5) +
  scale_fill_manual(values = c("#bae4bc", "#0868ac")) +
  labs(title = "Indicator differences across time and space") +
  plotTheme() + theme(legend.position="bottom")

```

# Kable Table

```{r kable_table}
kable(census_data_summary) %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 1.0")
```


# Pop Analysis

## Population Analysis

```{r population_analysis}

half_mile_buffer <- st_buffer(dcmetro, 2640) #Get half mile buffer

centroids <- st_centroid(all_census_data) #Get centroids of all census tracts

#Spatially join centroid of census tracts to buffers around metro stops and sum total population within 1/2 milf of buffer

metro_pop <- st_join(half_mile_buffer,centroids %>% dplyr::select(GEOID,year,total_pop,median_rent)) %>%  
  group_by(NAME,year) %>% summarize(pop = sum(total_pop),rent=mean(median_rent, na.rm=TRUE)) %>%
  st_drop_geometry() 

dcmetro <- inner_join(dcmetro,metro_pop,by='NAME')
```

## Make Population Map

```{r make_pop_map}

ggplot()+
  geom_sf(data = all_census_data)+
  geom_sf(data = dcmetro, color='blue', aes(size=pop))+
  geom_sf(data = tod_dissolve,fill='transparent',color='red',linewidth=0.5)+
  scale_size(range = c(0.1, 3))+
  mapTheme()+
  facet_wrap(~year)
```

## Make Rent Map

```{r make_rent_map}

ggplot()+
  geom_sf(data = all_census_data)+
  geom_sf(data = dcmetro, color='blue', aes(size=rent))+
  geom_sf(data = tod_dissolve,fill='transparent',color='red',linewidth=0.5)+
  scale_size(range = c(0.01, 4))+
  mapTheme()+
  facet_wrap(~year)
```

# Multi Ring Buffer Analysis


```{r make_rent_map}
allTracts_17 <- rbind(non_tod2017,tod2017)
allTracts_09 <- rbind(non_tod2009, tod2009)

MRB <- multipleRingBuffer(st_union(dcmetro), 27000, 1800)

rings_17 <- st_join(st_centroid(allTracts_17), MRB, join = st_intersects) %>% 
            st_drop_geometry() %>%           
            left_join(allTracts_17, by = "GEOID") %>%   
            mutate(distance_miles = distance / 5240) %>%            
            st_sf()

rings_09 <- st_join(st_centroid(allTracts_09), MRB, join = st_intersects) %>%  
            st_drop_geometry() %>%          
            left_join(allTracts_09, by = "GEOID") %>%            
            mutate(distance_miles = distance / 5240) %>%
            st_sf()

rings <- rbind(rings_17, rings_09)

ggplot()+  
         geom_sf(data = rings, aes(fill = distance_miles)) + 
         geom_sf(data = dcmetro, color = "white")  + 
         scale_fill_viridis_c() +mapTheme()

ggplot() +   
         geom_line(data = rings, aes(x = distance_miles, y = median_rent.x, color = year.y), stat = "summary", fun = mean) #add line tod vs non tod and how many houses there are tod vs non tod
```
